### 2024-05-19 thesis/input_files/analysis.tex
added conc sec
May 14th 2024
May 18th 2024

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the electricity load time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependet variable. Doing so, RMSE went down to 10.29. Forecast is visualized in \ref{fig:prophet_price2}, moreover, figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet v2 prediction}
    \label{fig:prophet_price2}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.1.png}
    \caption{Seasonalities breakdown}
    \label{fig:prophet_price2.1}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.2.png}
    \caption{Prices trend}
    \label{fig:prophet_price2.2}
\end{figure}
Prophet alongside with the right features is a good model in the context of electricity price forecasting, it is capcable of catching the trend, the peak and the lows.

@@

\begin{table}[!ht]
    \caption{Root mean squared errors}
    \label{tab:point_RMSE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
            Method/RMSE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 30.5892 & 29.0347 & 75.1682 & 63.1376 & 42.1156 & 36.6233 & 39.2496 & 38.8286 & 47.1307 & 68.0494 & 61.6005 & 30.3751 & 34.5523 & 33.4287 & 33.5581 \\
            TBATS & 15.0886 & 31.9720 & 99.1260 & 84.4890 & 51.9740 & 34.9055 & 18.4445 & 38.3246 & 74.1117 & 98.6600 & 84.3050 & 38.8433 & 16.6080 & 28.9902 & 41.6194 \\
            Prophet & 10.2936 & 14.4358 & 38.7551 & 63.4787 & 19.7474 & 17.5065 & 12.6926 & 14.2665 & 17.5466 & 23.5944 & 43.6666 & 20.8637 & 16.9493 & 19.1626 & 23.3889 \\
            KNN & 12.5429 & 11.6699 & 24.5057 & 18.3310 & 13.1821 & \textbf{11.9238} & 12.0044 & 14.5165 & 16.3132 & 15.1831 & 37.6457 & 16.4690 & 12.0324 & 11.3102 & 14.1717 \\
            SVR & 23.2421 & 25.8525 & 78.8043 & 67.6209 & 42.3748 & 32.7845 & 30.5971 & 35.3660 & 55.4213 & 77.9660 & 68.5799 & 30.2451 & 27.0345 & 28.3652 & 32.1480 \\
            LSTM & 9.1782 & 12.0669 & \textbf{22.7048} & \textbf{16.2087} & 16.0964 & 12.7936 & \textbf{10.8559} & 14.6173 & 19.7303 & 18.0200 & 43.2051 & 17.1856 & 10.3106 & 12.1347 & 17.5849 \\
            KrnRidge & 9.8619 & 11.6101 & 37.7824 & 35.2459 & \textbf{12.5160} & 14.5911 & 12.8791 & 17.4385 & 16.1131 & 17.1938 & 37.6961 & 14.0076 & 9.8441 & 10.7491 & 13.2975 \\
            KrnSVR & \textbf{9.1028} & \textbf{11.3117} & 22.9281 & 19.2132 & 12.6331 & 12.6018 & 11.3537 & \textbf{12.9506} & \textbf{14.9731} & \textbf{11.7765} & \textbf{37.3797} & \textbf{13.2694} & 
            \textbf{8.8522} & \textbf{10.6185} & \textbf{13.2602} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}
\end{table}

The MAE scores are contained in table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score.
\begin{table}[!ht]
    \caption{Mean absolute errors}
    \label{tab:point_MAE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
             Method/MAE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 27.7219 & 24.3823 & 65.4016 & 52.3873 & 34.8465 & 31.2663 & 35.9420 & 34.7316 & 37.0273 & 54.7075 & 48.9327 & 26.1975 & 31.8720 & 29.2306 & 28.3428 \\
            TBATS & 10.5408 & 23.8417 & 91.3153 & 74.5734 & 40.3258 & 26.0092 & 14.2141 & 24.7818 & 62.0452 & 86.7125 & 72.6912 & 28.4235 & 11.1184 & 21.1038 & 31.8555 \\
            Prophet & 8.6139 & 11.4307 & 28.3509 & 46.4480 & 15.6290 & 14.1278 & 10.1574 & 11.2360 & 14.0072 & 18.9565 & 27.1142 & 16.9678 & 14.6615 & 16.1750 & 18.3043 \\
            KNN & 9.4003 & 9.1126 & 19.6213 & 14.4906 & 10.6461 & \textbf{9.4354} & 8.7441 & 10.0065 & 12.4714 & 11.2006 & 20.3087 & 12.5416 & 9.2558 & 8.5937 & 10.5451 \\
            SVR & 20.2443 & 21.2526 & 69.3704 & 57.1107 & 34.3651 & 27.5875 & 27.4651 & 29.4204 & 43.7009 & 64.3522 & 55.4307 & 25.1080 & 24.0294 & 24.4049 & 26.5869 \\
            LSTM & 7.5217 & 9.4767 & \textbf{17.9851} & \textbf{13.2158} & 11.8118 & 9.8607 & \textbf{8.0561} & 10.8480 & 15.1008 & 13.8108 & 25.9906 & 13.3588 & 8.1526 & 9.8000 & 14.3186 \\
            KrnRidge & 7.5965 & 9.2130 & 28.2484 & 24.2366 & \textbf{10.0589} & 10.7758 & 9.5788 & 11.2166 & 12.3427 & 12.3306 & 20.0213 & 10.8434 & 7.1625 & \textbf{7.9710} & 10.0767 \\
            KrnSVR & \textbf{6.9581} & \textbf{8.6989} & 18.7739 & 15.1398 & 10.1395 & 9.5927 & 8.5731 & \textbf{9.0033} & \textbf{11.0618} & \textbf{8.5486} & \textbf{19.7383} & \textbf{10.5545} & \textbf{7.0395} & 8.1926 & \textbf{9.8125} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}            
\end{table}

Finally table \ref{tab:point_MAPE} reports the MAPE score for each method. What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
\begin{table}[!ht]
    \caption{Mean absolute percentage errors}
    \label{tab:point_MAPE}
    \begin{adjustbox}{width=\textwidth, height=3cm}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
             Method/MAPE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 0.2470 & 0.1935 & 0.3030 & 0.2601 & 0.2474 & 0.2667 & 0.3498 & 0.3058 & 0.1954 & 0.2432 & 0.4234 & 0.2114 & 0.2926 & 0.2502 & 0.2145 \\
            TBATS & 0.0833 & 0.1650 & 0.4285 & 0.3731 & 0.2482 & 0.1849 & 0.1234 & 0.1634 & 0.3181 & 0.4021 & 0.4880 & 0.1761 & 0.0905 & 0.1555 & 0.2058 \\
            Prophet & 0.0737 & 0.0854 & 0.1324 & 0.2453 & 0.1134 & 0.1169 & 0.0938 & 0.0912 & 0.0880 & 0.0992 & 0.3757 & 0.1389 & 0.1343 & 0.1356 & 0.1370 \\
            KNN & 0.0778 & 0.0682 & 0.0924 & 0.0761 & 0.0751 & 0.0740 & 0.0784 & 0.0744 & 0.0725 & 0.0555 & \textbf{0.3115} & 0.0888 & 0.0816 & 0.0699 & 0.0765 \\
            SVR & 0.1808 & 0.1640 & 0.3211 & 0.2829 & 0.2317 & 0.2260 & 0.2676 & 0.2451 & 0.2215 & 0.2867 & 0.4354 & 0.1870 & 0.2216 & 0.2044 & 0.1933 \\
            LSTM & 0.0642 & 0.0706 & 0.0896 & \textbf{0.0722} & 0.0798 & 0.0772 & \textbf{0.0710} & 0.0854 & 0.0909 & 0.0741 & 0.3578 & 0.0976 & 0.0719 & 0.0778 & 0.1037 \\
            KrnRidge & 0.0625 & 0.0685 & 0.1281 & 0.1207 & 0.0718 & 0.0797 & 0.0812 & 0.0760 & 0.0696 & 0.0587 & 0.3127 & 0.0779 & \textbf{0.0591} & \textbf{0.0637} & 0.0722 \\
            KrnSVR & \textbf{0.0568} & \textbf{0.0642} & \textbf{0.0892} & 0.0788 & \textbf{0.0714} & \textbf{0.0724} & 0.0740 & \textbf{0.0665} & \textbf{0.0629} & \textbf{0.0428} & 0.3122 & \textbf{0.0781} & 0.0607 & 0.0646 & \textbf{0.0713} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}
\end{table}

\section{Probabilistic forecasting}
In this subsection we compare performance of quantile regressors.
\\
- experiment for quantile estimator on gefcom2014 and we are good

>>>>

\subsection{Trigonometric seasonality Box-Cox transformation ARMA errors trend and seasonal components}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the electricity load time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats \cite{de2011forecasting} model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile its RMSE is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependet variable. Doing so, RMSE went down to 10.29. Forecast is visualized in \ref{fig:prophet_price2}, moreover.
% figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet v2 prediction}
    \label{fig:prophet_price2}
\end{figure}

% \begin{figure}[!h]
%     \includegraphics[width=\textwidth]{images/prophet_price2.1.png}
%     \caption{Seasonalities breakdown}
%     \label{fig:prophet_price2.1}
% \end{figure}

% \begin{figure}[!h]
%     \includegraphics[width=\textwidth]{images/prophet_price2.2.png}
%     \caption{Prices trend}
%     \label{fig:prophet_price2.2}
% \end{figure}
Prophet alongside with the right features is a good model in the context of electricity price forecasting, it is capcable of catching the trend, the peak and the lows.

@@

\begin{table}[!ht]
    \caption{Root mean squared errors}
    \label{tab:point_RMSE}
    \begin{adjustbox}{width=\textwidth}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
            Method/RMSE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 30.5892 & 29.0347 & 75.1682 & 63.1376 & 42.1156 & 36.6233 & 39.2496 & 38.8286 & 47.1307 & 68.0494 & 61.6005 & 30.3751 & 34.5523 & 33.4287 & 33.5581 \\
            TBATS & 15.0886 & 31.9720 & 99.1260 & 84.4890 & 51.9740 & 34.9055 & 18.4445 & 38.3246 & 74.1117 & 98.6600 & 84.3050 & 38.8433 & 16.6080 & 28.9902 & 41.6194 \\
            Prophet & 10.2936 & 14.4358 & 38.7551 & 63.4787 & 19.7474 & 17.5065 & 12.6926 & 14.2665 & 17.5466 & 23.5944 & 43.6666 & 20.8637 & 16.9493 & 19.1626 & 23.3889 \\
            KNN & 12.5429 & 11.6699 & 24.5057 & 18.3310 & 13.1821 & \textbf{11.9238} & 12.0044 & 14.5165 & 16.3132 & 15.1831 & 37.6457 & 16.4690 & 12.0324 & 11.3102 & 14.1717 \\
            SVR & 23.2421 & 25.8525 & 78.8043 & 67.6209 & 42.3748 & 32.7845 & 30.5971 & 35.3660 & 55.4213 & 77.9660 & 68.5799 & 30.2451 & 27.0345 & 28.3652 & 32.1480 \\
            LSTM & 9.1782 & 12.0669 & \textbf{22.7048} & \textbf{16.2087} & 16.0964 & 12.7936 & \textbf{10.8559} & 14.6173 & 19.7303 & 18.0200 & 43.2051 & 17.1856 & 10.3106 & 12.1347 & 17.5849 \\
            KrnRidge & 9.8619 & 11.6101 & 37.7824 & 35.2459 & \textbf{12.5160} & 14.5911 & 12.8791 & 17.4385 & 16.1131 & 17.1938 & 37.6961 & 14.0076 & 9.8441 & 10.7491 & 13.2975 \\
            KrnSVR & \textbf{9.1028} & \textbf{11.3117} & 22.9281 & 19.2132 & 12.6331 & 12.6018 & 11.3537 & \textbf{12.9506} & \textbf{14.9731} & \textbf{11.7765} & \textbf{37.3797} & \textbf{13.2694} & 
            \textbf{8.8522} & \textbf{10.6185} & \textbf{13.2602} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}
\end{table}

The MAE scores are contained in table \ref{tab:point_MAE}. Similarly to above, we can conclude that kernel methods stand out for achieving also the lowest MAE score.
\begin{table}[!ht]
    \caption{Mean absolute errors}
    \label{tab:point_MAE}
    \begin{adjustbox}{width=\textwidth}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
             Method/MAE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 27.7219 & 24.3823 & 65.4016 & 52.3873 & 34.8465 & 31.2663 & 35.9420 & 34.7316 & 37.0273 & 54.7075 & 48.9327 & 26.1975 & 31.8720 & 29.2306 & 28.3428 \\
            TBATS & 10.5408 & 23.8417 & 91.3153 & 74.5734 & 40.3258 & 26.0092 & 14.2141 & 24.7818 & 62.0452 & 86.7125 & 72.6912 & 28.4235 & 11.1184 & 21.1038 & 31.8555 \\
            Prophet & 8.6139 & 11.4307 & 28.3509 & 46.4480 & 15.6290 & 14.1278 & 10.1574 & 11.2360 & 14.0072 & 18.9565 & 27.1142 & 16.9678 & 14.6615 & 16.1750 & 18.3043 \\
            KNN & 9.4003 & 9.1126 & 19.6213 & 14.4906 & 10.6461 & \textbf{9.4354} & 8.7441 & 10.0065 & 12.4714 & 11.2006 & 20.3087 & 12.5416 & 9.2558 & 8.5937 & 10.5451 \\
            SVR & 20.2443 & 21.2526 & 69.3704 & 57.1107 & 34.3651 & 27.5875 & 27.4651 & 29.4204 & 43.7009 & 64.3522 & 55.4307 & 25.1080 & 24.0294 & 24.4049 & 26.5869 \\
            LSTM & 7.5217 & 9.4767 & \textbf{17.9851} & \textbf{13.2158} & 11.8118 & 9.8607 & \textbf{8.0561} & 10.8480 & 15.1008 & 13.8108 & 25.9906 & 13.3588 & 8.1526 & 9.8000 & 14.3186 \\
            KrnRidge & 7.5965 & 9.2130 & 28.2484 & 24.2366 & \textbf{10.0589} & 10.7758 & 9.5788 & 11.2166 & 12.3427 & 12.3306 & 20.0213 & 10.8434 & 7.1625 & \textbf{7.9710} & 10.0767 \\
            KrnSVR & \textbf{6.9581} & \textbf{8.6989} & 18.7739 & 15.1398 & 10.1395 & 9.5927 & 8.5731 & \textbf{9.0033} & \textbf{11.0618} & \textbf{8.5486} & \textbf{19.7383} & \textbf{10.5545} & \textbf{7.0395} & 8.1926 & \textbf{9.8125} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}            
\end{table}

Finally table \ref{tab:point_MAPE} reports the MAPE score for each method. What can be concluded is that kernel regressors are also the best in terms of MAPE score; with KrnSVR being the top one.
\begin{table}[!ht]
    \caption{Mean absolute percentage errors}
    \label{tab:point_MAPE}
    \begin{adjustbox}{width=\textwidth}
        \begin{tabular}{lrrrrrrrrrrrrrrr}
            \toprule
             Method/MAPE & Task 1 & Task 2 & Task 3 & Task 4 & Task 5 & Task 6 & Task 7 & Task 8 & Task 9 & Task 10 & Task 11 & Task 12 & Task 13 & Task 14 & Task 15 \\
            \midrule
            MLR & 0.2470 & 0.1935 & 0.3030 & 0.2601 & 0.2474 & 0.2667 & 0.3498 & 0.3058 & 0.1954 & 0.2432 & 0.4234 & 0.2114 & 0.2926 & 0.2502 & 0.2145 \\
            TBATS & 0.0833 & 0.1650 & 0.4285 & 0.3731 & 0.2482 & 0.1849 & 0.1234 & 0.1634 & 0.3181 & 0.4021 & 0.4880 & 0.1761 & 0.0905 & 0.1555 & 0.2058 \\
            Prophet & 0.0737 & 0.0854 & 0.1324 & 0.2453 & 0.1134 & 0.1169 & 0.0938 & 0.0912 & 0.0880 & 0.0992 & 0.3757 & 0.1389 & 0.1343 & 0.1356 & 0.1370 \\
            KNN & 0.0778 & 0.0682 & 0.0924 & 0.0761 & 0.0751 & 0.0740 & 0.0784 & 0.0744 & 0.0725 & 0.0555 & \textbf{0.3115} & 0.0888 & 0.0816 & 0.0699 & 0.0765 \\
            SVR & 0.1808 & 0.1640 & 0.3211 & 0.2829 & 0.2317 & 0.2260 & 0.2676 & 0.2451 & 0.2215 & 0.2867 & 0.4354 & 0.1870 & 0.2216 & 0.2044 & 0.1933 \\
            LSTM & 0.0642 & 0.0706 & 0.0896 & \textbf{0.0722} & 0.0798 & 0.0772 & \textbf{0.0710} & 0.0854 & 0.0909 & 0.0741 & 0.3578 & 0.0976 & 0.0719 & 0.0778 & 0.1037 \\
            KrnRidge & 0.0625 & 0.0685 & 0.1281 & 0.1207 & 0.0718 & 0.0797 & 0.0812 & 0.0760 & 0.0696 & 0.0587 & 0.3127 & 0.0779 & \textbf{0.0591} & \textbf{0.0637} & 0.0722 \\
            KrnSVR & \textbf{0.0568} & \textbf{0.0642} & \textbf{0.0892} & 0.0788 & \textbf{0.0714} & \textbf{0.0724} & 0.0740 & \textbf{0.0665} & \textbf{0.0629} & \textbf{0.0428} & 0.3122 & \textbf{0.0781} & 0.0607 & 0.0646 & \textbf{0.0713} \\
            \bottomrule
            \end{tabular}            
    \end{adjustbox}
\end{table}

\section{Probabilistic forecasting}
% In this subsection we compare performance of quantile regressors.
% \\
% - experiment for quantile estimator on gefcom2014 and we are good

\section{Conclusions}
